---
title: R Basics seq(), rep(), and standard deviation with basic functions
author: Matthew Haines
date: '2017-09-14'
slug: r-basics-seq-rep-and-standard-deviation-with-basic-functions
categories:
  - R
tags:
  - Functions
  - Statistics
---
# R Series 3

##Functions seq(),rep(), and standard deviation
In the last post we went over significant functions like mean() and diff() this post will examine the seq() or sequence function, the rep() or repeat function, and a way to calculate standard deviation with basic functions. 

First, we will examine the sequence function. The sequence function runs by inputing two numbers seperated by a comma in the parenthesis. You then have the option of inputing parameters. For example, if we wanted a list of all odd numbers from 1-100 the function would look like this.

```{r}
seq(1,100, by=2)
```

Now, let's look at the repeat function. The repeat function does exactly what it says, it repeats the input you give it. So pretty basic example below:

```{r}
rep("z", 7)
```

Now that we have the simple stuff out of the way, let's look at standard deviation. Standard deviation is calculated from the deviations of each data point from the mean of the data set. Then, if you square these results and get the mean of those values you have calcualated the data set's variance. From there all you have to do is take the square root of the variance. An example of these steps are listed below.

```{r}
Exdata<- c(4,3,2,11,29,54,59,35,41,46,25,47,60,54,46,49,46,41,34,22,6)
Exmean<- (sum(Exdata)/length(Exdata))

Exmean
mean(Exdata)

Exsum<- (Exdata - Exmean)
Exsum

Exsumsq<- (Exdata - Exmean)^2
Exsumsq

Exvar<- (sum(Exsumsq)/length(Exsumsq)-1)
Exvar

ExStdev<- (Exvar)^0.5
ExStdev
```

So let's recap that equation a bit. We used the combine function to make the data set. Then, the sum function to get the sum of all the values in the data set and the length function to calculate a mean for the data set. Next, I ran the mean function to double check the work. The next line of code outlines how to calculate the difference of each individual point from the mean and it is stored into the Exsum variable. Then, we squared that data set to get ready to calculate the set's variance. Then, we calculated the set's variance and finally took the square root of that set to finally find the set's standard deviation. For those of you wondering, the example data is the number of home runs Babe Ruth hit in each of his seasons in major league baseball. That's it for this post check out more for more info on R. 